---
title: "generate_plots"
output:
  html_document: default
  pdf_document: default
date: "2023-02-27"
---
# load packages

```{r}
suppressWarnings({
library(ggplot2)
library(tidyverse)
library(data.table)
library("plotrix") 
library("patchwork")
library(purrr)
library(ggpubr)
library(stringi)
library(dplyr)
library(MCL)
})
rm(list=ls())
```



# set up plot dir
```{r}
plots_dir<- file.path(getwd(),"plots")
if(!dir.exists(plots_dir)) dir.create(plots_dir, recursive = TRUE, showWarnings = FALSE)
```



# read data from csv
```{r}
data23m <- read.csv(file = file.path("..", "data", "sl_23_morning","monstera.csv")) 
data23a <- read.csv(file = file.path("..", "data", "sl_23_afternoon","monstera.csv")) 
data24m <- read.csv(file = file.path("..", "data", "sl_24_morning","monstera.csv")) 
data24a <- read.csv(file = file.path("..", "data", "sl_24_afternoon","monstera.csv")) 

data <- rbind(data23m, data23a, data24m, data24a)
```



# data wrangling
## check and subset NA
```{r}
data$selection <- na_if(data$selection, '')
na_number <- xtabs(~is.na(data$selection))[2]
na_ratio <- paste(round(100*(na_number/nrow(data)), 2), "%", sep="")
print(paste("There are",na_number,"responses marked as NA, makes",na_ratio,"of total.", sep = ' '))
data <- data %>% filter(!is.na(data$selection))

```


## encode responses, translating indices of selections to strings 
```{r}
response_type <- c("ra","ew","oh")
df <- data %>% mutate(selection = as.numeric(stri_sub(selection,-1,-1))) %>%
  mutate(continutations_indices = gsub("\\[|\\]|, " ,"",continutations_indices)) %>%
  mutate(response_index = as.numeric(substr(continutations_indices,selection,selection))) %>%
  mutate(response = response_type[response_index])
```

## aggregate data
```{r}
df <- subset(df, conditions != "dummy")
data_high <- subset(df, expectancy == "high")
data_medium <- subset(df, expectancy == "medium")

aggregated_data <- df %>%
  group_by(response,particle,expectancy) %>% 
  summarise(n = n())
```

# relative freqeuncies within high and medium expectancy
```{r}
(tm_medium <- prop.table(table(data_medium$particle, data_medium$response), margin=1))
(tm_high <- prop.table(table(data_high$particle, data_high$response), margin=1))
```

# Perform Clustering using MCL
```{r}
mcl(tm_medium, addLoops = FALSE, allow1 = TRUE)
mcl(tm_high, addLoops = FALSE, allow1 = TRUE)
```

# plot data
## generate plot
```{r}
my_theme <- function() {
  theme_minimal() +
    theme(
      plot.title = element_text(size = 20, face = "bold"),
      axis.title.x = element_text(size = 20, margin = margin(t = 10, r = 0, b = 0, l = 0)),
      axis.title.y = element_text(size = 20, margin = margin(t = 0, r = 10, b = 0, l = 0)),
      axis.text.x = element_text(size = 20, angle = 45, hjust = 1),
      axis.text.y = element_text(size = 18),
      legend.title = element_text(size = 20),
      legend.text = element_text(size = 18),
      strip.text.x = element_text(size = 20),
      strip.text.y = element_text(size = 18, angle = 0),
      panel.grid.major = element_line(colour = "grey90", size = 0.25),
      panel.grid.minor = element_blank(),
      panel.border = element_blank(),
      panel.background = element_blank(),
      legend.position="bottom"
    )
}

p <- ggplot(aggregated_data,aes(x=particle,y=n,fill=response)) + 
    geom_bar(stat="identity",position=position_dodge()) +
    facet_grid(~expectancy) +
    scale_fill_brewer(palette="Paired") + my_theme() + ylab("Abs. Freq. of responses")

p
```
## save plot to pdf
```{r}
pdf(file = file.path(plots_dir,"mean_freq_facet_without_dummy_sl.pdf"))
print(p)
while (!is.null(dev.list()))
dev.off()
```

xtabs(~response+particle+expectancy, data=df)


## re-encode responses EW vs . other and plot
```{r}
#enocode new binary response variable
df$response_ew_other <- ifelse(df$response == "ew", "ew", "other")
df$response_ew_other <- as.factor(df$response_ew_other)
df$response_ew_other_num <- ifelse(df$response_ew_other == "ew", 1, 0)
#dat$response_ew_other

aggregated_data <- df %>%
  group_by(response_ew_other,particle,expectancy) %>% 
  summarise(n = n())

p_ew_other <- ggplot(aggregated_data,aes(x=particle,y=n,fill=response_ew_other)) + 
    geom_bar(stat="identity",position=position_dodge()) +
    facet_grid(~expectancy) +
    scale_fill_brewer(palette="Paired") + my_theme() + ylab("Abs. Freq. of responses") #+
#    scale_fill_discrete(name="response")

p_ew_other

pdf(file = file.path(plots_dir,"mean_freq_facet_without_dummy_ew_other_sl.pdf"))
#print(p_ew_other)
p_ew_other
while (!is.null(dev.list()))
dev.off()
```

## re-encode responses RA vs. OH and plot

```{r}
df_oh_ra <- subset(df, response %in% c("ra","oh"))
#dat$response_ew_other

aggregated_data_oh_ra <- df_oh_ra %>%
  group_by(response,particle,expectancy) %>% 
  summarise(n = n())


p_oh_ra <- ggplot(aggregated_data_oh_ra,aes(x=particle,y=n,fill=response)) + 
    geom_bar(stat="identity",position=position_dodge()) +
    facet_grid(~expectancy) +
    scale_fill_brewer(palette="Paired") + my_theme() + ylab("Abs. Freq. of responses") +
    scale_fill_discrete(name="response")

p_oh_ra

png(file = file.path(plots_dir,"mean_freq_facet_without_dummy_oh_ra_sl.png "))
print(p_oh_ra)
while (!is.null(dev.list()))
dev.off()
```